Fusing Filters with Integer Linear Programming

### 2014-05-12 2014betterfusionforfilters/section/03-SizeInference.tex
Add notes to ILP section
transducers

\subsection{Transducers}
Unlike previous work, we do allow combinators with different iteration sizes to be fused together.
For example, an operation on filtered data may be fused with the filter operation that generates the data, even though the iteration sizes are different.
More generally, if the output size of a combinator is different to its iteration size, it is a \emph{transducer} from the iteration size to the output size.
As usual, a transducer may fuse with other nodes of the same iteration size,
but transducers may also fused with nodes with iteration size the same as the transducer's output size.
For our set of combinators, the only transducer is @filter@.

\TODO{Explain that a transducer of filter is the filter's parent. so $trans(trans(n)) \not= n$}

>>>>

% -----------------------------------------------------------------------------
\subsection{Transducers}
Unlike previous work, we do allow combinators with different iteration sizes to be fused together.
For example, an operation on filtered data may be fused with the filter operation that generates the data, even though the iteration sizes are different.

We define the concept of \emph{transducers} as combinators having a different output size to their iteration size.
As usual, a transducer may fuse with other nodes of the same iteration size, but transducers may also fuse with nodes having iteration size the same as the transducer's output size.
For our set of combinators, the only transducer is @filter@.

Looking back at the @normalize2@ example, the iteration size of @gts@ and @sum1@ are both $k_{xs}$.
The iteration size of @sum2@ is $k_{gts}$, and @gts@ is a transducer from $k_{xs}$ to $k_{gts}$.
Despite it not being known whether $k_{gts} = k_{xs}$, the three nodes @gts@, @sum1@ and @sum2@ can all be fused together.
The two nodes @sum2@ and @gts@ can be fused together, since the output size of @gts@ is the iteration size of @sum2@.
Similarly, @sum1@ and @gts@ can be fused together, as they have the same iteration size.
\begin{code}
 normalize2 xs
  = let sum1 = fold   (+)  0   xs
        gts  = filter (> 0)    xs
        sum2 = fold   (+)  0   gts
        ...
\end{code}

% \ben{Do other systems use the concept of transducers?} \TODO{Explain that a transducer of filter is the filter's parent. so $trans(trans(n)) \not= n$}
% No,
We now define a function $trans$, to find the parent transducer of a given combinator.
